\documentclass[12pt]{report}

 % Add this line to your preamble
% Language setting
% Replace `English' with e.g. `Spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size

\usepackage[top=3cm,bottom=3cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
	
% Load the setspace package
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{float}
\usepackage{filecontents}
\usepackage{pgfplotstable}

\usepgfplotslibrary{fillbetween}
\usetikzlibrary{arrows.meta, bending}
\pgfplotsset{compat=newest}
\usetikzlibrary{decorations.markings}
\lstdefinestyle{mystyle}{
    commentstyle=\color{gray},
    keywordstyle=\color{blue},
    numberstyle=\tiny\color{gray},
    stringstyle=\color{purple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

% Useful packages
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{tikz}
\usepackage{natbib}
\pgfplotsset{width=12cm,compat=1.9}
\setstretch{1.5}

\bibliographystyle{unsrtnat}

\usetikzlibrary{shapes.arrows, positioning, arrows.meta}


\title{Financial Frictions and the Cleansing Effects of Recessions}
\author{Riccardo Dal Cero}


\begin{document}
\section{Free debt case}

\begin{align*}
    \begin{cases}
        V(k_0) = U(d_0^*) + \beta V(k_1), \\
        k_1 = \left[ f(k_{0}) + (1 - \delta) k_{0}  - d^*_0 \right] , \\
        d^*_0 = [\beta V'(k_{1})]^{-1},\\
        k_0 \text{ given.}
    \end{cases}
\end{align*}
The welfare function instead of being log-linear as in the thesis I employ the following quadratic preferences:

\begin{equation}
   W_0= \sum_{t=0}^{\infty}\beta^t\left(\epsilon + \gamma d_t - \frac{\eta}{2}d_t^2 \right)
\end{equation}

where \(\gamma\) and \(\eta\) are assumed to be positive. Moreover in order to be solvable is necessary to approximate
the intertemporal constraint with a linear quadratic approximation.
The steady-state equations are:

\begin{align}
    \frac{1}{\beta}=\alpha Z \hat{k}^{\alpha -1}+1-\delta \label{eq1}\\ 
    \widehat{d} =Z \hat{k}^{\alpha}-\delta \hat{k}\label{eqds}
\end{align}

Using the Taylor theorem to the log utility function:
\begin{equation}
    \ln(d_t) \approx \ln(\hat{d}) + \frac{1}{\hat{d}}(d_t - \hat{d}) - \frac{1}{2\hat{d}^2}(d_t - \hat{d})^2
\end{equation}


Doing the same for the flow of funds constraint:

\begin{align*}
    k_{t+1}= Z \hat{k}^{\alpha}+(1-\delta)\hat{k}+\alpha Z \hat{k}^{\alpha-1}(k_t-\hat{k})+(1-\delta)(k_t-\hat{k})-\hat{d}-(d_t-\hat{d}),\\
    k_{t+1}= Z \hat{k}^{\alpha} - \hat{d}+\left(\alpha Z \hat{k}^{\alpha-1} +1 -\delta\right)(k_t-\hat{k})-(d_t-\hat{d}),
\end{align*}

Using the equation (\ref{eq1}, \ref{eqds}), we get the new :

\begin{align*}
    k_t-\hat{k}=\frac{1}{\beta}(k_t-\hat{k})-(d_t-\hat{d})
\end{align*}

the firm problem in the absence of debt:
\begin{align*}
    W_0= \left(\epsilon + \gamma d_t - \frac{\eta}{2}d_t^2 + \beta V(k_1)\right),
    s.t. k_1 = (1+r)k_0-d_0
\end{align*}
where k and d are deviation from the mean and \(\left(1+r\right)=\frac{1}{\beta}\)

The first order condition is: 
\begin{equation}
    \lambda - \eta d_0^* = \beta V^{\prime}(k_1)
\end{equation}
from the dynamic constraint we obtain \(d_0=(1+r)k_0-k_1\), substituting into the belman equation, we get:
\begin{align*}
    \begin{cases}
        V_(k_0)= \epsilon + \gamma [(1+r)k_0 -k_1] - \frac{\eta}{2}\left[\left(1+r\right)k_0 -k_1\right]^2 +\beta V(k_1)\\
        \gamma -\eta\left[\left(1+r\right)k_0-k_1\right]=\beta V^{\prime}(k_1), \\ 
        k_0 given
    \end{cases}
\end{align*}
Our guess is similar to the quadratic utility function 
\begin{equation}
    V(k_t)=g+h k_t = \frac{m}{2}k_t^2
\end{equation}
Exploiting the guess and the second equation of the above system we get:
\begin{equation}
    k_0=\frac{1}{1+r}\left[\frac{\gamma-\beta h}{\eta}+\frac{\eta-\beta m}{\eta} k_1\right] .
\end{equation}
Substituting the guess into \(V(k_0), V(k_1)\), we get:
\begin{equation}
    \left\{\begin{array}{c}
    g+\frac{h}{1+r}\left(\frac{\gamma-\beta h}{\eta}\right)+\frac{m}{2(1+r)^2}\left(\frac{\gamma-\beta h}{\eta}\right)^2=\varepsilon+\frac{\gamma}{\eta}(\gamma-\beta h)-\frac{(\gamma-\beta h)^2}{2 \eta}+\beta g \\
    \frac{h}{1+r}\left(\frac{\eta-\beta m}{\eta}\right)+\frac{m}{(1+r)^2}\left(\frac{\eta-\beta m}{\eta}\right)\left(\frac{\gamma-\beta h}{\eta}\right)=-\frac{\gamma \beta m}{\eta}+\frac{(\gamma-\beta h) \beta m}{\eta}+\beta h \\
    \frac{m}{2(1+r)^2}\left(\frac{\eta-\beta m}{\eta}\right)^2=-\frac{(\beta m)^2}{2 \eta}+\frac{\beta m}{2}
    \end{array}\right.
\end{equation}
the solutions of the system are:

\begin{align*}
    \begin{cases}
        m=\frac{\eta}{\beta}\left[1-\beta(1+r)^2\right] \\
        h=\frac{\gamma}{\beta} \frac{\left[\beta(1+r)^2-1\right]}{r} \\ 
        g=\frac{\epsilon}{1-\beta}+\frac{\gamma^2}{2 \eta \beta(1-\beta)}\left[\frac{1-\beta(1+r)}{r}\right]^2 \\
    \end{cases}
\end{align*}
Therefore the optimal path for capital is:
\begin{align*}
    k_1 = \frac{k_0 (1 + r) - \frac{\gamma \left( \frac{r - \beta (1 + r)^2 + 1}{r} \right)}{\eta}}{\beta (1 + r)^2}
\end{align*}
Given the substitution \( 1 + r = \frac{1}{\beta} \), we can further simplify the equation for \( k_1 \).

1. Substitute \( 1 + r = \frac{1}{\beta} \) into the expressions for \( m \) and \( h \):

\[ m = \frac{\eta}{\beta} \left[ 1 - \beta \left(\frac{1}{\beta}\right)^2 \right] = \frac{\eta}{\beta} \left[ 1 -
\frac{1}{\beta^2} \right] = \frac{\eta}{\beta} \left[ \frac{\beta^2 - 1}{\beta^2} \right] = \frac{\eta (\beta^2 -
1)}{\beta^3} \]


\[ h = \frac{\gamma}{\beta} \frac{\left[ \beta \left(\frac{1}{\beta}\right)^2 - 1 \right]}{r} = \frac{\gamma}{\beta}
\frac{\left[ \frac{1}{\beta} - 1 \right]}{r} = \frac{\gamma}{\beta} \frac{\left[ 1 - \beta \right]}{r \beta} =
\frac{\gamma (1 - \beta)}{r \beta^2} \]


2. Substitute these values into the equation for \( k_1 \):

The original equation is:

\[ k_1 = \frac{k_0 (1 + r) - \frac{\gamma - \beta h}{\eta}}{1 - \frac{\beta m}{\eta}} \]

Substituting the new values of \( m \) and \( h \):

\[ \gamma - \beta h = \gamma - \beta \left( \frac{\gamma (1 - \beta)}{r \beta^2} \right) = \gamma - \frac{\gamma (1 -
\beta)}{r \beta} = \gamma \left( 1 - \frac{1 - \beta}{r \beta} \right) \]


Given \( r = \frac{1}{\beta} - 1 \):

\[ \gamma \left( 1 - \frac{1 - \beta}{r \beta} \right) = \gamma \left( 1 - \frac{1 - \beta}{(\frac{1}{\beta} - 1) \beta} \right) \]

\[ = \gamma \left( 1 - \frac{1 - \beta}{\frac{1 - \beta \beta}{\beta}} \right) = \gamma \left( 1 - \frac{1 - \beta}{1 -
\beta} \right) = \gamma \left( 1 - 1 \right) = 0 \]


Thus, we see that \( \gamma - \beta h = 0 \).

Now substituting \( m \) and simplifying the denominator:

\[ \frac{\beta m}{\eta} = \frac{\beta \left( \frac{\eta (\beta^2 - 1)}{\beta^3} \right)}{\eta} = \frac{\beta (\beta^2 -
1)}{\beta^3} = \frac{\beta^2 - 1}{\beta^2} \]


\[ 1 - \frac{\beta m}{\eta} = 1 - \frac{\beta^2 - 1}{\beta^2} = \frac{\beta^2 - (\beta^2 - 1)}{\beta^2} =
\frac{1}{\beta^2} \]


Therefore:

\[ k_1 = \frac{k_0 (1 + r) - 0}{\frac{1}{\beta^2}} = k_0 (1 + r) \beta^2 \]

Since \( 1 + r = \frac{1}{\beta} \):

\[ k^{\star}_1 = k_0 \left( \frac{1}{\beta} \right) \beta^2 = \beta k_0 \]

This is the optimal path for capital.
The policy function is:
\[d^{\star}_0=\frac{1}{\beta} k_0 - k_0 \beta = \frac{1 - \beta^2}{\beta} k_0\]
Reminding that \(\beta\) in the steady state of dividends is equal to \((\alpha Z \hat{k}^{\alpha -1}+1-\delta)^{-1}\),
therefore we can rewrite the optimal path of capital as: 
\[k_1 = \beta k_0 = \frac{1}{\alpha Z \hat{k}^{\alpha -1}+1-\delta} k_0 \]
Thus a deviation from the steady state of capital \(k_0\) either positive or negative will follow a converging path to
the steady state since \(\beta\) is less than 1.
The optimal path of dividends is:
\begin{equation}
    d^{\star} = \frac{1-\left(\alpha Z \hat{k}^{\alpha-1}+1-\delta\right)^{-2}}{\left(\alpha Z \hat{k}^{\alpha-1}+1-\delta\right)^{-1}} k_0
\end{equation}
This result suggests that a positive shock to capital will lead to a positive shock to dividends, but the magnitude of
the shock will be smaller than the shock to capital, since \(0<(1-\beta^2)/\beta<1\).
\section{simulation}
Let's use the following parameters:
\begin{table}[H]
    \centering
    \begin{tabular}{lcc}
    \hline Parameter & Symbol & Value \\
    \hline \hline
    Depreciation rate & $\delta$ & 0.07 \\
    Returns to scale & $\alpha$ & 0.70 \\
    Firm productivity & $Z$ & 1 \\
    Discount factor & $\beta$ & 0.95 \\
    \hline
    \end{tabular}
    \caption{Benchmark calibration}
\end{table}

The following plot shows the evolution of capital and dividends over 100 periods, with log deviations from the steady state.

\pgfplotstableread[col sep=comma]{output_data/quadratic_function.csv}\datatable

\begin{figure}[h!]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            width=12cm,
            height=8cm,
            grid=major,
            xlabel={Periods},
            ylabel={Log Deviations from Steady State},
            legend pos=north east,
            title={Evolution of Capital and Dividends Over Time},
            legend style={font=\small},
            ymajorgrids=true,
            xmajorgrids=true,
        ]
        \addplot[
            color=blue,
        ] table[x=Period,y={Capital (log deviations)}] {\datatable};
        \addlegendentry{Capital (log deviations)}

        \addplot[
            color=green,
        ] table[x=Period,y={Dividends (log deviations)}] {\datatable};
        \addlegendentry{Dividends (log deviations)}
        \end{axis}
    \end{tikzpicture}
    \caption{Log deviations from steady state for capital and dividends over 100 periods.}
    \label{fig:impulse_response1}
\end{figure}
 
The impulse response function \ref{fig:impulse_response1} represents the effect of a positive $10\%$ shock to the steady state level of capital (blue)
and on dividends (green). The rate of convergence is a function of \(\beta\). Lets for example consider the same
positive shock on capital but with a discount factor of 0.9:
\pgfplotstableread[col sep=comma]{output_data/quadratic_function_lowerb.csv}\datatable

\begin{figure}[!h]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            width=12cm,
            height=8cm,
            grid=major,
            xlabel={Periods},
            ylabel={Log Deviations from Steady State},
            legend pos=north east,
            title={Evolution of Capital and Dividends Over Time},
            legend style={font=\small},
            ymajorgrids=true,
            xmajorgrids=true,
        ]
        \addplot[
            color=blue,
        ] table[x=Period,y={Capital (log deviations)}] {\datatable};
        \addlegendentry{Capital (log deviations)}

        \addplot[
            color=green,
        ] table[x=Period,y={Dividends (log deviations)}] {\datatable};
        \addlegendentry{Dividends (log deviations)}
        \end{axis}
    \end{tikzpicture}
    \caption{Log deviations from steady state for capital and dividends over 100 periods with a discount factor of 0.9.}
    \label{fig:impulse_response2}
\end{figure}
As the impulse response function \ref{fig:impulse_response2} shows, the rate of convergence is faster with a lower discount rate. Therefore,
an increase in the risk free rate will lead to a faster convergence to the steady state level of capital and dividends,
after a positive shock to capital.
Regarding productivity shocks, they affected only the steady state level of capital and dividends, but not the rate of convergence.

\section{Leverage Case}

In the previous section, the firm was unable to finance investment through debt. In this section, we will consider the
case in which the firm can borrow from the financial market. In this model, firms face a borrowing constraint and a
participation constraint of the financial intermediaries.  

The participation constraint of the financial intermediaries is defined by the opportunity cost of lending to the firm,
which is the risk-free rate, and is represented by the following equation: 
\begin{equation}
    R_t \cdot b_t p + (1-p) \mu f(k_t) = R_f b_t \label{eq:partecipation}
\end{equation}
where \(R_t\) is the return of the firm, \(b_t\) is the amount of debt, \(p\) is the probability of default, \(\mu\) is
the monitoring cost, \(f(k_t)\) is the return of the firm in the case of no default, and \(R_f\) is the risk-free rate.  

The borrowing constraint is defined by the following equation:
\begin{equation}
    b_t \leq l k_t \quad 0 < l < 1 \label{eq:borrowing}
\end{equation}
where \(l\) is the leverage ratio. Both constraints are binding conditions for simplicity, meaning that if a firm fails
to meet one of the two constraints, the firm will default. 
\subsection{The firm problem}
The firm's objective is to maximize:

\[
\max_{{\{d_{t}\}}_{t=0}^{+\infty}}V_0 = \sum_{t=0}^{+\infty}{\beta^t U(d_t)}
\]

subject to:
\begin{enumerate}
    \item the flow of funds constraint: \(I_t + R b_{t} + d_t = f(k_t) + b_{t+1}\)
    \item the investment function \(I_t = k_{t+1} - k_{t}\left(1-\delta\right)\) 
    \item the financing constraint \(b_t=l k_t\)
    \item the participation constraint of lenders \(R_t=\frac{R_f}{p}  -\frac{ 1-p }{ p }\frac{\mu f(k_t)}{b_t}\)
\end{enumerate}
Consolidating and linearizing the constraints, we get:
\begin{align}
    \tilde{k}_{t+1} &= \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p} \frac{ \tilde{k}_t}{1-l} - \frac{\tilde{d}_t}{1-l} \label{eq:linearized_flow}
\end{align}
where \(\tilde{k}_t\) and \(\tilde{d}_t\) are deviations from the steady state. The calculations are shown in the
appendix.
The utility function is the same as in the previous section. Therefore, the problem can be rewritten as the following system:
\begin{align*}
     \begin{cases}
         V(k_0) = U(d_0^*) + \beta V(k_1), \\
         k_1 =  a \lambda \tilde{k}_0 - \lambda \tilde{d}^*_0, \\
         k_0 \text{ given.}
     \end{cases}
 \end{align*}
where \(\lambda = \frac{1}{1-l}\) and \(a = \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p}\).

The value function \( V(k_0) \) is:
\[
V(k_0) = A \ln(k_0) + B,
\]
where:
\[
A = \frac{1 - \beta}{\beta \ln(a \lambda)},
\]
\[
B = \frac{\ln(d_0^*)}{1 - \beta}.
\]

The optimal consumption \( d_0^* \) maximizes:
\[
\ln(d_0^*) + \beta V(a \lambda k_0 - \lambda d_0^*).
\]

The first-order condition gives:
\[
d_0^* = \frac{a \lambda k_0}{1 + \beta \lambda}.
\]

The optimal path for capital is:
\[
k_{t+1} = a \lambda k_t - \lambda d_t^*,
\]
substituting \( d_t^* \):
\[
k_{t+1} = k_t \left( \frac{a \lambda (1 + (\beta - 1) \lambda)}{1 + \beta \lambda} \right).
\]

This provides the policy function for optimal consumption \( d_0^* \) and the optimal evolution path for capital \( k_t \). See the appendix for the complete solution.

  \section{Simulation}

This section presents the results of our dynamic programming model with linear utility. We simulate the impulse response functions of capital and consumption to a 10\% increase in productivity \(Z\).

\begin{table}[H]
    \centering
    \begin{tabular}{lcc}
    \hline Parameter & Symbol & Value \\
    \hline \hline
    Depreciation rate & $\delta$ & 0.07 \\
    Returns to scale & $\alpha$ & 0.70 \\
    Firm productivity & $Z$ & 1 \\
    Discount factor & $\beta$ & 0.95 \\
    Risk-free rate & $R_f$ & 0.03 \\
    Initial capital & $k_0$ & 1 \\
    Initial $\lambda$ & $\lambda_{\text{initial}}$ & 0.1 \\
    \hline
    \end{tabular}
    \caption{Benchmark calibration}
    \label{tab:benchmark_calibration}
\end{table}

The simulation data is imported from a CSV file and visualized using charts. The data includes the paths of capital and consumption under the baseline scenario and the scenario where productivity \(Z\) is increased by 10\%.

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            width=\textwidth,
            height=0.5\textwidth,
            xlabel={Period},
            ylabel={Capital (K)},
            legend pos=outer north east,
            title={Impulse Response of Capital}
        ]
        \addplot table [x=Period, y=Capital, col sep=comma] {output_data/impulse_response.csv};
        \addlegendentry{Capital (K)}
        \addplot table [x=Period, y=Capital10, col sep=comma] {output_data/impulse_response.csv};
        \addlegendentry{Capital (K) with Z increased by 10\%}
        \end{axis}
    \end{tikzpicture}
    \caption{Impulse Response of Capital}
    \label{fig:impulse_response_capital}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            width=\textwidth,
            height=0.5\textwidth,
            xlabel={Period},
            ylabel={Consumption (d)},
            legend pos=outer north east,
            title={Impulse Response of Consumption}
        ]
        \addplot table [x=Period, y=Consumption, col sep=comma] {output_data/impulse_response.csv};
        \addlegendentry{Consumption (d)}
        \addplot table [x=Period, y=Consumption10, col sep=comma] {output_data/impulse_response.csv};
        \addlegendentry{Consumption (d) with Z increased by 10\%}
        \end{axis}
    \end{tikzpicture}
    \caption{Impulse Response of Consumption}
    \label{fig:impulse_response_consumption}
\end{figure}

The charts above show the impulse response functions for capital and consumption over 50 periods. The baseline scenario is represented by solid lines, while the scenario with a 10\% increase in productivity \(Z\) is shown with dashed lines.

\begin{itemize}
    \item \textbf{Capital (K):} The left chart shows the response of capital to a 10\% increase in productivity. The increase in productivity leads to a higher capital accumulation over time.
    \item \textbf{Consumption (d):} The right chart shows the response of consumption to a 10\% increase in productivity. Higher productivity also results in increased consumption over time.
\end{itemize}

\appendix

\section{appendix}
\subsection{A: Linearization of Constraints}
\label{sec:linearization}
As in the previous section, we will linearize the constraints around the steady state. By consolidating constraints
\ref{eq:partecipation}, \ref{eq:borrowing}, and the flow of funds, we obtain: 
\begin{equation}
    k_{t+1} = \left[ \frac{p + \mu - \mu p}{p}f(k_{t}) + \frac{p - \delta p - R_f l}{p} k_{t} - d_t \right] \cdot \left(1-l\right)^{-1} \label{eq:flow}
\end{equation}

First, we compute the steady state of the model (using the results from the master thesis):

\begin{align}
    &Z \alpha \hat{k}^{\alpha-1} = \frac{p - p l - \beta p + \beta \delta p + \beta R_f l}{\beta \left(p + \mu - \mu p\right)} \label{eq:k_opt_path} \\
    &\widehat{d} =\frac{p+\mu-\mu p}{p}f(\hat{k})-\left(\frac{l R_f+\delta p - l p}{p}\right)\hat{k} \label{eq:div_opt_path}
\end{align}

Using the Taylor theorem for the log utility function, we get:

\begin{align*}
    \ln(d_t) = \ln(\hat{d}) + \frac{1}{\hat{d}}(d_t - \hat{d}) - \frac{1}{2\hat{d}^2}(d_t - \hat{d})^2
\end{align*}

To approximate the following constraint using a Taylor approximation up to the first degree:
\begin{equation}
    k_{t+1} = \left[ \frac{p + \mu - \mu p}{p} f(k_{t}) + \frac{p - \delta p - R_f l}{p} k_{t} - d_{t} \right] \cdot \left(1-l\right)^{-1} \label{eq:flow}
\end{equation}

We linearize around the steady state values given by:
\begin{align}
    Z \alpha \hat{k}^{\alpha-1} &= \frac{p - p l - \beta p + \beta \delta p + \beta R_f l}{\beta \left(p + \mu - \mu p\right)} \label{eq:k_opt_path} \\
    \hat{d} &= \frac{p + \mu - \mu p}{p} f(\hat{k}) - \left(\frac{l R_f + \delta p - l p}{p}\right) \hat{k} \label{eq:div_opt_path}
\end{align}

First, define the function:
\begin{equation}
g(k_t, d_t) = \left[ \frac{p + \mu - \mu p}{p} f(k_{t}) + \frac{p - \delta p - R_f l}{p} k_{t} - d_{t} \right] \cdot \left(1-l\right)^{-1}
\end{equation}

We approximate \( g(k_t, d_t) \) around \( \hat{k} \) and \( \hat{d} \) using a first-order Taylor expansion:
\begin{equation}
g(k_t, d_t) \approx g(\hat{k}, \hat{d}) + \left. \frac{\partial g}{\partial k_t} \right|_{k_t = \hat{k}} (k_t - \hat{k}) + \left. \frac{\partial g}{\partial d_t} \right|_{d_t = \hat{d}} (d_t - \hat{d})
\end{equation}

Calculate the partial derivatives:
\begin{equation}
\frac{\partial g}{\partial k_t} = \left(1-l\right)^{-1} \left[ \frac{p + \mu - \mu p}{p} f'(k_t) + \frac{p - \delta p - R_f l}{p} \right]
\end{equation}

\begin{equation}
\frac{\partial g}{\partial d_t} = -\left(1-l\right)^{-1}
\end{equation}

Using \( f(k_t) = Z k_t^\alpha \) and \( f'(\hat{k}) = Z \alpha \hat{k}^{\alpha-1} \), we get:
\begin{equation}
\left. \frac{\partial g}{\partial k_t} \right|_{k_t = \hat{k}} = \left(1-l\right)^{-1} \left[ \frac{p + \mu - \mu p}{p} Z \alpha \hat{k}^{\alpha-1} + \frac{p - \delta p - R_f l}{p} \right]
\end{equation}

Substitute the steady-state value:
\begin{equation}
Z \alpha \hat{k}^{\alpha-1} = \frac{p - p l - \beta p + \beta \delta p + \beta R_f l}{\beta \left(p + \mu - \mu p\right)}
\end{equation}

So,
\begin{equation}
\left. \frac{\partial g}{\partial k_t} \right|_{k_t = \hat{k}} = \left(1-l\right)^{-1} \left[ \frac{p + \mu - \mu p}{p} \cdot \frac{p - p l - \beta p + \beta \delta p + \beta R_f l}{\beta \left(p + \mu - \mu p\right)} + \frac{p - \delta p - R_f l}{p} \right]
\end{equation}

Simplify:
\begin{equation}
\left. \frac{\partial g}{\partial k_t} \right|_{k_t = \hat{k}} = \left(1-l\right)^{-1} \left[ \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p} \right]
\end{equation}

Thus, the linearized form is:
\begin{align}
    k_{t+1} &\approx \hat{k} + \frac{k_t - \hat{k}}{1-l}\left[ \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p} \right] \nonumber 
     - \frac{d_t - \hat{d}}{1-l}
\end{align}
taking the steady state values of \(k_t\) on the lhs, we get:
\begin{align}
    k_{t+1} -\hat{k} &= \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p} \frac{ k_t - \hat{k}}{1-l} \nonumber 
    - \frac{d_t - \hat{d}}{1-l}
\end{align}
Defining the deviation from the steady state as:
\begin{align}
    \begin{cases}
        k_t = \hat{k} + \tilde{k}_t \\
        d_t = \hat{d} + \tilde{d}_t
    \end{cases}
\end{align}
we get:
\begin{align}
    \tilde{k}_{t+1} &= \frac{p - p l - \beta p + \beta \delta p + \beta R_f l + p - \delta p - R_f l}{\beta p} \frac{ \tilde{k}_t}{1-l} - \frac{\tilde{d}_t}{1-l} \label{eq:linearized_flow}
\end{align}
The policy function is:
  
\begin{equation}
d_0^* = \frac{a \lambda k_0}{1 + \beta \lambda}
\end{equation}

The optimal path for capital is:

\begin{equation}
    k_{t+1} = k_t \left( \frac{a \lambda (1 + (\beta - 1) \lambda)}{1 + \beta \lambda} \right)
\end{equation}



\subsection{B: Resolution with Log Utility and Policy Function}

  Consider the utility function \( U(d_t) = \ln(d_t) \). The Bellman equation is:
  
  \[
  V(k_0) = \ln(d_0^*) + \beta V(k_1),
  \]
  
  where \( k_1 = a \lambda k_0 - \lambda d_0^* \) and \( k_0 \) is given.
  
  Assume the value function \( V(k) \) takes the form:
  
  \[
  V(k) = A \ln(k) + B,
  \]
  
  where \( A \) and \( B \) are constants to be determined.
  
  Substitute the hypothesized value function into the Bellman equation:
  
  \[
  A \ln(k_0) + B = \ln(d_0^*) + \beta (A \ln(a \lambda k_0) + \beta B).
  \]
  
  Using \( \ln(a \lambda k_0) = \ln(a \lambda) + \ln(k_0) \):
  
  \[
  A \ln(k_0) + B = \ln(d_0^*) + \beta A \ln(a \lambda) + \beta A \ln(k_0) + \beta B.
  \]
  
  Matching coefficients gives:
  
  \[
  A (1 - \beta) = \beta A \ln(a \lambda),
  \]
  
  \[
  B (1 - \beta) = \ln(d_0^*).
  \]
  
  Solving for \( A \) and \( B \):
  
  \[
  a \lambda = e^{\frac{1 - \beta}{\beta}},
  \]
  
  \[
  B = \frac{\ln(d_0^*)}{1 - \beta}.
  \]
  
  The value function \( V(k_0) \) is:
  
  \[
  V(k_0) = A \ln(k_0) + B,
  \]
  
  where:
  
  \[
  A = \frac{1 - \beta}{\beta \ln(a \lambda)},
  \]
  
  \[
  B = \frac{\ln(d_0^*)}{1 - \beta}.
  \]
  
  The optimal consumption \( d_0^* \) maximizes:
  
  \[
  \ln(d_0^*) + \beta V(a \lambda k_0 - \lambda d_0^*).
  \]
  
  The first-order condition gives:
  
  \[
  d_0^* = \frac{a \lambda k_0}{1 + \beta \lambda}.
  \]
  
  The optimal path for capital is:
  
  \[
  k_{t+1} = a \lambda k_t - \lambda d_t^*,
  \]
  
  substituting \( d_t^* \):
  
  \[
  k_{t+1} = k_t \left( \frac{a \lambda (1 + (\beta - 1) \lambda)}{1 + \beta \lambda} \right).
  \]
  
  This provides the policy function for optimal consumption \( d_0^* \) and the optimal evolution path for capital \( k_t \).
  
 



\end{document}